SDSS-IV MaNGA: The Universal Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-07-17 maintex.tex
minor edits

This star formation enhancement increases with closer projected separations between the two galaxies \citep{Ellison:2008, Scudder:2012} but has still shown to be detectable out to projected separations of 150 kpc \citep{Patton:2013}. The level of the star formation enhancement also increases with smaller stellar mass ratios, $|\Delta {\rm log} {\it M} |$, between the two galaxies \citep{Ellison:2008}. Similar mass galaxies show the strongest sSFR effects while large central galaxies interacting with small satellite galaxies show weaker sSFR effects. Because of this, most surveys will split their sample into major mergers, $|\Delta$log($M$)$|$ $\le$ 0.5, and minor mergers, 0.5 $\le$ $|\Delta$log($M$)$|$ $\le$ 1.0, where major mergers are expected to show the strongest interaction induced effects.

@@

%%%Purpose%%%
In our previous work, \citet{Fu:2018}, we built a sample of paired galaxies in the MaNGA survey where both components of the pair were contained within field of view of a single integral field unit (IFU). We found that approximately 5.7\% of the MaNGA galaxies have a companion galaxy contained within the field-of-view of a single IFU. In this work, we will use a version of this sample updated for the recent MaNGA Data Release 16 (DR16) and we will further supplement this with a sample of companion galaxies identified outside the field-of-view of the MaNGA IFUs.

@@

The MaNGA survey targets galaxies from a subset of 41,154 galaxies from the NASA-Sloan Atlas (NSA) with a redshift range of 0.01 $<$ z $<$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M_{\rm i}}$ $<$ -24.0, where $\mathcal{M_{\rm i}}$ is the rest frame i-band magnitude within the survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 \reff\ and 2.5 \reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the internal data release, MPL-8, which covers 6142 unique galaxies.

@@

We use our catalog of photometric objects and the \citet{Simard:2011} catalog to mask foregournd/background objects which may contaminate the data. Objects in our photometric catalog are masked out with a 2\arcsec\ circular aperture and the galaxies from \citet{Simard:2011} are masked out with a 2.0 \reff\ elliptical aperture, using the r-band S\'ersic apertures from the catalog.

@@

In Figures \ref{fig:ssfr_prof} and \ref{fig:ssfr_mass}, we show the profile of $\Delta$log(sSFR) as a function of galactocentric radius between two different methods of comparing paired galaxies to control galaxies. In the first method (Section \ref{sec:mass-bin}) where pair and control galaxies are compared within different stellar mass bins. The difference profiles from this method show a central enhancement to the log(sSFR) of $\sim$0.3 $\pm$ 0.1 dex. The enhancement linearly falls to zero at 1.0 \reff\ after which the disks of the paired galaxies have zero enhancement/suppression to their sSFR.

The $\Delta$log(sSFR) profile following the second method shows the same central sSFR enhancement albeit with smaller variances. Within the inner 1.0 \reff, there is a variance between the profile of 0.1 dex. Beyond 1.0 \reff, the variance increases 0.2 dex. This indicates that the second method of pair$-$ control comparison is superior to the first method.

@@

We show the mass growth rate profile for each of the four mass ranges in Figure \ref{fig:mass_gain_sum}. The mass growth rate is largely restricted to the centers of the galaxies and falls to zero around 1.0 \reff, following profile of the $\Delta$log(sSFR). The mass growth rate is strongly dependent on the stellar mass of the paired galaxies.

@@

In Figure \ref{fig:prof_comp} we look at the $\Delta$log(sSFR) profile as a function of galactocentric radius between our work and previous works. \citet{Pan:2019} examines a set of paired galaxies in the MaNGA survey while \citet{Thorp:2019} examines a set of post-merger galaxies in the MaNGA survey. The galaxy pairs in \citet{Pan:2019} have sSFR enhancements that are roughtly $\sim$0.05$-$0.2 dex higher than our sample. The $\Delta$log(sSFR) profile is closest to our work at the centers of the galaxy pairs but then begins to diverge further out into the galaxies' disks. Our paired galaxies show zero sSFR enhancement in their disks past \reff\ $=$ 1.0 while the galaxy pairs in \citet{Pan:2019} show an enhancement of $\sim$0.15 dex to the sSFR in their disks.

The post-merger galaxies in \citet{Thorp:2019} have sSFR enhancements between $\sim$0.05$-$0.15 higher than our sample of paired galaxies. Their $\Delta$log(sSFR) profile follows our sample's profile closely between 0.25$-$1.0 \reff. The centers of the post-merger galaxies feature a greater level of sSFR enhancement in their centers compared to our galaxies pairs of $\Delta$log(sSFR) $=$ 0.5, roughly 0.25 dex higher than our paired galaxies. The post-merger galaxies also feature enhanced sSFR in their disk past 1.0 \reff\ of $\sim$0.1$-$0.15 dex. The heightened level of sSFR enhancement in the centers of post-merger galaxies in comparison to merging galaxies seems consistent with the idea that a second and greater burst of star formation occurs as the two merging galaxies coalesce into a single galaxy.

@@

\begin{enumerate}
\item The sSFR is centrally enhanced within the inner 1.0 \reff\ by 0.30 $\pm$ 0.1 dex. 
\item The disks of the paired galaxies, R/\reff\ $=$ 1.0$-$2.5, show no signs of either sSFR enhancement or sSFR suppression. 
\item The sSFR enhancement in paired galaxies is independent of the total stellar mass of the paired galaxies, Massive galaxies experience the same level of sSFR enhancement as low mass galaxies.
\item The sSFR enhancement in paired galaxies does depend on the ratio between the masses of the paired galaxies. Galaxies with small mass ratios, $|\Delta$log(M)$|$ $=$ 0.0$-$0.5, see an enhancement to sSFR of 0.4 $\pm$ 0.1 dex, which is 0.15 dex higher than the average paired galaxies in the survey. We have also showed that paired galaxies with large mass ratios, 1.0 $<$ $|\Delta$log(M)$|$ $<$ 2.0, still show substantial levels of enhancement to the sSFR, 0.1$-$0.2 dex. 
\item Further, in paired galaxies with large mass ratios, the less massive galaxies experience higher levels of sSFR enhancement than the more massive galaxy in the pair, $\sim$0.1 dex.
\item The sSFR enhancement increases with closer projected separations where galaxies within 5$-$15 kpc have their sSFR enhanced by 0.5$-$0.6 $\pm$ 0.1 dex. The galaxies within 5 kpc experience no central enhancement to their sSFR. We believe that these galaxies have just passed their first pericenter and the merger induced gas-inflow have yet to reach the centers of these galaxies.
\item The paired galaxies will experience a faster mass growth rate of secular evolution of 1.6$-$2.3$\times$. The growth rate is highest for massive galaxies; however, the the low mass galaxies will see a higher fractional mass increase in comparison to the preexisting stellar mass.

>>>>

This star formation enhancement increases with closer projected separations between the two galaxies \citep{Ellison:2008, Scudder:2012} but has still shown to be detectable out to projected separations of 150 kpc \citep{Patton:2013}. The level of the star formation enhancement also increases with smaller stellar mass ratios, $|\Delta$log($M$)$|$, between the two galaxies \citep{Ellison:2008}. Similar mass galaxies show the strongest sSFR effects while large central galaxies interacting with small satellite galaxies show weaker sSFR effects. Because of this, most surveys will split their sample into major mergers, $|\Delta$log($M$)$|$ $\le$ 0.5, and minor mergers, 0.5 $\le$ $|\Delta$log($M$)$|$ $\le$ 1.0, where major mergers are expected to show the strongest interaction induced effects.

@@

%%%Purpose%%%
In our previous work, \citet{Fu:2018}, we built a sample of paired galaxies in the MaNGA survey where both components of the pair were contained within field of view of a single integral field unit (IFU). We found that approximately 5.7\% of the MaNGA galaxies have a companion galaxy contained within the field-of-view of a single IFU. In this work, we will use a version of this sample updated for the recent MaNGA Data Release 16 (DR16) \citep{Ahumada:2020} and we will further supplement this with a sample of companion galaxies identified outside the field-of-view of the MaNGA IFUs.

@@

The MaNGA survey targets galaxies from a subset of 41,154 galaxies from the NASA-Sloan Atlas (NSA) with a redshift range of 0.01 $<$ z $<$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M_{\rm i}}$ $<$ -24.0, where $\mathcal{M_{\rm i}}$ is the rest frame i-band magnitude within the survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 R/\reff\ and 2.5 R/\reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the internal data release, MPL-8, which covers 6142 unique galaxies.

@@

We use our catalog of photometric objects and the \citet{Simard:2011} catalog to mask foregournd/background objects which may contaminate the data. Objects in our photometric catalog are masked out with a 2\arcsec\ circular aperture and the galaxies from \citet{Simard:2011} are masked out with a 2.0 R/\reff\ elliptical aperture, using the r-band S\'ersic apertures from the catalog.

@@

In Figures \ref{fig:ssfr_prof} and \ref{fig:ssfr_mass}, we show the profile of $\Delta$log(sSFR) as a function of galactocentric radius between two different methods of comparing paired galaxies to control galaxies. In the first method (Section \ref{sec:mass-bin}) where pair and control galaxies are compared within different stellar mass bins. The difference profiles from this method show a central enhancement to the log(sSFR) of $\sim$0.3 $\pm$ 0.1 dex. The enhancement linearly falls to zero at 1.0 R/\reff\ after which the disks of the paired galaxies have zero enhancement/suppression to their sSFR.

The $\Delta$log(sSFR) profile following the second method shows the same central sSFR enhancement albeit with smaller variances. Within the inner 1.0 R/\reff, there is a variance between the profile of 0.1 dex. Beyond 1.0 R/\reff, the variance increases 0.2 dex. This indicates that the second method of pair$-$ control comparison is superior to the first method.

@@

We show the mass growth rate profile for each of the four mass ranges in Figure \ref{fig:mass_gain_sum}. The mass growth rate is largely restricted to the centers of the galaxies and falls to zero around 1.0 R/\reff, following profile of the $\Delta$log(sSFR). The mass growth rate is strongly dependent on the stellar mass of the paired galaxies.

@@

In Figure \ref{fig:prof_comp} we look at the $\Delta$log(sSFR) profile as a function of galactocentric radius between our work and previous works. \citet{Pan:2019} examines a set of paired galaxies in the MaNGA survey while \citet{Thorp:2019} examines a set of post-merger galaxies in the MaNGA survey. The galaxy pairs in \citet{Pan:2019} have sSFR enhancements that are roughtly $\sim$0.05$-$0.2 dex higher than our sample. The $\Delta$log(sSFR) profile is closest to our work at the centers of the galaxy pairs but then begins to diverge further out into the galaxies' disks. Our paired galaxies show zero sSFR enhancement in their disks past 1.0 R/\reff\ while the galaxy pairs in \citet{Pan:2019} show an enhancement of $\sim$0.15 dex to the sSFR in their disks.

The post-merger galaxies in \citet{Thorp:2019} have sSFR enhancements between $\sim$0.05$-$0.15 higher than our sample of paired galaxies. Their $\Delta$log(sSFR) profile follows our sample's profile closely between 0.25$-$1.0 R/\reff. The centers of the post-merger galaxies feature a greater level of sSFR enhancement in their centers compared to our galaxies pairs of $\Delta$log(sSFR) $=$ 0.5, roughly 0.25 dex higher than our paired galaxies. The post-merger galaxies also feature enhanced sSFR in their disk past 1.0 R/\reff\ of $\sim$0.1$-$0.15 dex. The heightened level of sSFR enhancement in the centers of post-merger galaxies in comparison to merging galaxies seems consistent with the idea that a second and greater burst of star formation occurs as the two merging galaxies coalesce into a single galaxy.

@@

\begin{enumerate}
\item The sSFR is centrally enhanced within the inner 1.0 R/\reff\ by 0.30 $\pm$ 0.1 dex. 
\item The disks of the paired galaxies, R/\reff\ $=$ 1.0$-$2.5, show no signs of either sSFR enhancement or sSFR suppression. 
\item The sSFR enhancement in paired galaxies is independent of the total stellar mass of the paired galaxies, Massive galaxies experience the same level of sSFR enhancement as low mass galaxies.
\item The sSFR enhancement in paired galaxies does depend on the ratio between the masses of the paired galaxies. Galaxies with small mass ratios, $|\Delta$log(M)$|$ $=$ 0.0$-$0.5, see an enhancement to sSFR of 0.4 $\pm$ 0.1 dex, which is 0.15 dex higher than the average paired galaxies in the survey. We have also showed that paired galaxies with large mass ratios, 1.0 $<$ $|\Delta$log(M)$|$ $<$ 2.0, still show substantial levels of enhancement to the sSFR, 0.1$-$0.2 dex. 
\item Further, in paired galaxies with large mass ratios, the less massive galaxies experience higher levels of sSFR enhancement than the more massive galaxy in the pair, $\sim$0.1 dex.
\item The sSFR enhancement increases with closer projected separations where galaxies within 5$-$15 kpc have their sSFR enhanced by 0.5$-$0.6 $\pm$ 0.1 dex. The galaxies within 5 kpc experience no central enhancement to their sSFR. We believe that these galaxies have just passed their first pericenter and the merger induced gas-inflow have yet to reach the centers of these galaxies.
\item The paired galaxies will experience a faster mass growth rate of secular evolution of 1.6$-$2.3$\times$. The growth rate is highest for massive galaxies; however, the the low mass galaxies will see a higher fractional mass increase in comparison to the preexisting stellar mass.
